`cesR`: A package to make using the Canadian Election Study a little easier

### 2020-08-27 paper.Rmd
updated headers

When called, the `get_ces()` function returns a requested CES survey as a data object and prints to the console the associated citation and URL for the survey dataset repository. The function takes one argument in the form of a character string. This argument is a vector member that has been associated with a CES survey through the body of code in the `get_ces()` function that when used calls the download URL for that survey on an associated GitHub repository named `ces_data`. If the provided character string argument matches a member of the built-in vector `ces_codes`, the associated file is downloaded using the `download.file()` function from the `utils` R package (R Core Team, 2020) as a compressed .zip folder and is stored temporarily in `inst/extdata` directory in the greater package directory. Upon downloading the file, the compressed folder is unzipped using the `unzip()` function from the `utils` R package (R Core Team, 2020) and read into R using either the `read_dta()` or `read_sav()` functions from the `haven` R package [@citehaven] depending on the file extension of the downloaded file. A data frame is then assigned using the `assign()` function  from the `base` R package (R Core Team, 2020) as a data object in the global environment. The downloaded file and file directory are then removed from the computer using the `unlink()` function from the `base` R package (R Core Team, 2020). Finally, the recommended citation for the requested survey dataset and URL of the survey data storage location are printed in the console (see [*2.1.1: `get_ces()` basics*](#211-get_ces-basics) for an example of a basic `get_ces()` function call).

If the provided character string argument does not have a match in the built-in vector, then the function process is stopped and a warning message stating `Error in get_ces(): Warning: Code not in table` is printed in the RStudio console (see [*2.1.2: `get_ces()` error*](#212-get_ces-error)).

#### `get_ces()` basics
```
# install the cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

#### `get_ces()` error
```
# install the cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

CES survey dataset files uses either a `.dta()` or `.sav()` file extension, meaning the datasets are loaded into R and RStudio as the type labelled. The `get_ces()` function does not convert the values of the loaded tables to a factor type so that personal workflow practices are not interfered with. It is suggested that to convert the dataset values to a factor type that the `to_factor()` function from the `labelled` package (Lamarange, 2020) be used (see [*2.1.3: `to_factor()` function*](#213-to_factor-function) for an example as to using the `to_factor()` function and `labelled` package).

#### `to_factor()` function
```
# install cesR package from GitHub
devtools::intall_github("hodgettsp/cesR")

@@

The `get_cescodes()` function provides a user a means of looking up the argument calls that are used to access each CES survey dataset. The `get_cescodes()` function does not take any arguments. Instead when the function is called it prints to the console a dataframe that contains the survey codes and their associated argument calls. See [*2.2.1: `get_cescodes()` function*](#22-example-1-get_cescodes-function) for an example of the print output and the [*3. Vignette*](#3-vignette) section for an example of the `get_cescodes()` used as part of the full `cesR` package.

#### `get_cescodes()` function
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

## `get_question()`
The `get_question()` provides users with the ability to look up a survey question associated with a given column name. The function takes two arguments in the form of character strings, those being the name of a data object and the name of a column in the given data object. The function works such that it checks whether the given data object exists using the `exists()` function from the `base` package (R Core Team, 2020). If the object does not exist, the function will print out a warning in the console stating `Warning: Data object does not exist` (see [*2.3.2: `get_question()` data object error*](#232-get_question-data-object-error) for an example of this error). If the object does exist, `get_question()` will check if the given column name exists in the given data object. This is done using a combination of the `hasName()` function from the `utils` package and the `get()` function from the `base` package (R Core Team, 2020). The `hasName()` function checks if the given column name is in the given data object. Because the arguments are given as character strings the `get()` function is used to return the actual data object instead of the provided character string. Otherwise, the `hasName()` function would only check if the given column name argument occurred in the given character string argument and not the actual data object. If the column does not exist in the data object a warning is printed in the console stating `Warning: Variable is not in dataset`. See [*2.3.3: `get_question()` variable error*](#233-get_question-variable-error) for an example of this error message. If the given column exists in the given data object, `get_question()` will print the variable label of the given column to the console using a combination of the `var_label()` function from the `labelled` package (Larmarange, 2020) and the `get()` function from the `base` package (R Core Team, 2020). See [*2.3.1: `get_question()` basics*](#231-get_question-basics) for an example of the general use of the `get_question()` function and the [*3. Vignette*](#3-vignette) section for its use as part of the whole `cesR` package.

#### `get_question()` basics
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

#### `get_question()` data object error
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

#### `get_question()` variable error
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

In addition to being usable with the data objects created from the [`get_ces()`](#21-get_ces) function, the `get_question()` function is structured in such a way that it can also be used to return the column label for the `decon` dataset or any dataset of the labelled type. [*2.3.4: `get_question()` `decon`*](#224-get_question-decon) presents an example of the `get_question()` function being used with the `decon` dataset.

#### `get_question()` `decon`
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

### `get_preview()`

To make the data exploratory phase easier it may be prudent to have access to a truncated version of a dataset. Additionally, a truncated dataset provides a resource that can be used by educators in the teaching of exploratory data analysis. The `get_preview()` function provides such truncated versions of the CES datasets. The function takes two arguments, a character string to call a survey of the same style as used for the [`get_ces()`](#21-get_ces) function and a numerical value that sets the number of rows returned (see [2.4.1: `get_preview()` basics](#241-get_preview-basics)). If no value is provided for the number of rows, a default of six is returned (see [2.4.2: `get_preview()` no value provided](#242-get_preview-no-value-provided)).

#### `get_preview()` basics
```
#install cesR package
devtools::install_github("hodgettsp/cesR")

@@

#### `get_preview()` no value provided
```
# install cesR package
devtools::install_github("hodgettsp/cesR")

@@

### get_decon()
When called, creates a subset of the 2019 CES online survey under the name `decon` (demographics and economics) that provides a tool for educators in the teaching of the analysis of large survey datasets. The `get_decon()` function takes no arguments. The function first checks the global environment if an object named `decon` exists using the `exists()` function from the `base` package (R Core Team, 2020). This prevents the `decon` dataset from being recreated if the object exists. If the `get_decon()` function is run when an object with the name `decon` already exists a warning will print in the console stating `Error in get_decon() : Warning: File already exists.` See [*2.5.1: `get_decon()` error*](#251-get_decon-error) for an example of the error message. If a situation arises in which the `decon` dataset needs to be recreated, then the best course of action is to use the `rm()` function from the `base` package (R Core Team, 2020) to remove the `decon` object and then run the `get_decon()` function again (see [*2.5.2: `get_decon()` reload*](#252-get_decon-reload)).

#### `get_decon()` error
```
# install cesR package
devtools::install_github("hodgettsp/cesR")

@@

#### `get_decon()` reload
```
# install cesR package
devtools::install_github("hodgettsp/cesR")

@@

After checking if an object with the name `decon` exists, the `get_decon()` function performs similarly to the `get_ces()` function. It assigns a temporary file extension with a .zip type using the `tempfile()` function from the `base` package, and calls on the URL for the 2019 CES online survey and temporarily downloads the associated file using the `download.file()` function from the `base` package (R Core Team, 2020). After unpacking the compressed folder with the `unzip()` function from the `utils` package (R Core Team, 2020), the file is read into R using the `read_dta()` function from the `haven` package (Wickham & Miller, 2020). Using the `select()` function from the `dplyr` package, 20 variables are selected from the main CES 2019 online survey and renamed using the `rename` function from the `dplyr` package (Wickham et al., 2020). The dataframe values are then converted to a factor type using the `to_factor()` function from the `labelled` package (Lamarange, 2020). A new variable, consisting of participants' responses to their self-perceived position on the political spectrum, is then created from two variables using the `unite()` function from the `tidyr` package (Wickham & Henry, 2020) and all empty cells are replaced with `NA` values using the `mutate()` function from the `dplyr` package (Wickham et al., 2020). The dataset is made available in the global environment by using the `assign()` function from the `base` package to create a new data object under the name `decon`. Lastly, the downloaded files are removed from the computer using the `unlink()` function from the `base` package (R Core Team, 2020) and a citation for the 2019 CES online survey and a URL to the storage location is printed to the console. See [*2.5.3: `get_decon()` basics*](#253-get_decon-basics) for an example of a general use of the `get_decon()` function.

#### `get_decon()` basics

@@

To begin, install and load the `cesR` package (and all other necessary packages) into RStudio. Currently, this is currently only available through the use of the `install_github` function from the `devtools` package (Wickham et al., 2020). During installation, RStudio may request to update other packages. It is best to press enter with an empty line (see [*3.1.1: Install `cesR`*](#311-install-cesR)).

#### Install `cesR`
```
# uncomment any package that needs to be installed
# install.packages("devtools")
# install.packages("labelled")
# install.packages("dplyr")
# install.packages("tidyr")

@@

Upon installation and loading of the `cesR` package, we can use the `get_cescodes()` function to look up the code for the desired CES survey. In this case that is the 2019 CES phone survey (see [*3.1.2: Lookup CES codes*](#312-lookup-ces-codes)).

#### Lookup CES codes
```
# look up survey codes
get_cescodes()
```
```
> get_cescodes()
   index ces_survey_code get_ces_call_char
1      1     ces2019_web     "ces2019_web"
2      2   ces2019_phone   "ces2019_phone"
3      3     ces2015_web     "ces2015_web"
4      4   ces2015_phone   "ces2015_phone"
5      5   ces2015_combo   "ces2015_combo"
6      6         ces2011         "ces2011"
7      7         ces2008         "ces2008"
8      8         ces2004         "ces2004"
9      9         ces0411         "ces0411"
10    10         ces0406         "ces0406"
11    11         ces2000         "ces2000"
12    12         ces1997         "ces1997"
13    13         ces1993         "ces1993"
14    14         ces1988         "ces1988"
15    15         ces1984         "ces1984"
16    16         ces1974         "ces1974"
17    17         ces7480         "ces7480"
18    18      ces72_jnjl      "ces72_jnjl"
19    19       ces72_sep       "ces72_sep"
20    20       ces72_nov       "ces72_nov"
21    21         ces1968         "ces1968"
22    22         ces1965         "ces1965"
```
In the printed data frame we can see that the argument code for the 2019 CES phone survey is "ces2019_phone" (see [*3.1.2: Lookup CES codes*](#312-lookup-ces-codes)). This code can be used with the `get_preview()` function to look up a truncated preview of the 2019 CES phone survey dataset (see [*3.1.3: Preview dataset*](#313-preview-dataset)).

#### Preview dataset
```
# get preview of first 8 rows of 2019 CES phone survey dataset
get_preview("ces2019_phone", 8)
```
 sample_id | survey_end_CES | survey_end_mont~ | survey_end_day_~ | num_attempts_CES | interviewer_id_~ | interviewer_gen~ | language_CES | phonetype_CES |
-----------|----------------|------------------|------------------|------------------|------------------|------------------|--------------|---------------|
<dbl>      | <chr>          | <dbl>            | <dbl>            |  <dbl>           |  <dbl>           |     <chr>        |    <fct>     | <fct>         |
18         | 2019-09-23 15~ |               9  |              23  |              5   |         161182   | Female           | (2) French   |  (2) Wireless |
32         | 2019-09-12 18~ |               9  |              12  |              1   |         151152   | Male             | (1) English  |  (2) Wireless |
39         | 2019-09-10 18~ |               9  |              10  |              1   |         161182   | Female           | (2) French   |  (2) Wireless |
59         | 2019-10-10 15~ |              10  |              10  |              6   |         147601   | Female           | (2) French   |  (2) Wireless |
61         | 2019-09-12 16~ |               9  |              12  |              1   |         151152   | Male             | (2) French   |  (2) Wireless |
69         | 2019-09-17 17~ |               9  |              17  |              1   |           2503   | Female           | (2) French   |  (2) Wireless |       
157        | 2019-09-12 16~ |               9  |              12  |              1   |           2529   | Male             |  (2) French  | (2) Wireless   | 
158        | 2019-09-14 10~ |                9 |              14  |               4  |         142515   | Male             |  (2) French  | (2) Wireless   |

Using the same survey code arguemtn with the `get_ces()` function we can retrieve the complete 2019 CES phone survey (see [*3.1.4: `get_ces()`*](#314-get_ces)). Additionally, we can see when `get_ces()` is called, the survey citation and link to repository prints to the console.

#### `get_ces()`
```
# call 2019 CES phone survey
get_ces("ces2019_phone")
```
```
TO CITE THIS SURVEY FILE: Stephenson, Laura B; Harell, Allison; Rubenson, Daniel; Loewen, Peter John, 2020, '2019 Canadian Election Study - Phone Survey', https://doi.org/10.7910/DVN/8RHLG1, Harvard Dataverse, V1, UNF:6:eyR28qaoYlHj9qwPWZmmVQ== [fileUNF]

@@

To check that the dataset has loaded into RStudio correctly, it is best to check the dataset with the `head()` function from the `utils` package (R Core Team, 2020). Unfortunately, the `head()` function does not work with labelled data, so it is best to convert the values to factors using the `to_factor()` function from the `labelled` package (Larmarange, 2020) (see [*3.1 Example 3: Check 2019 CES phone data*](#315-check-ces-phone-data)).

#### Check 2019 CES phone data
```
# convert values to factor type
ces2019_phone <- to_factor(ces2019_phone)
head(ces2019_phone)

@@

With the dataset loaded, the `select()` function from the `dplyr` package (Wickham et al., 2020) is an efficient way of selecting the required variable columns. In this case, the language, phone type, weight, citizenship, year of birth, interviewer gender, identified gender, province, satisfaction with Canadian democracy, and most important election issue are selected using the column index (see [*3.1.4: Select columns*](#314-select-columns)). Additionally, the columns are not all named in an understandable way and so will be renamed using the `rename()` function from the `dplyr` package (Wickham et al., 2020).

#### Select columns
```
ces2019_phone_subset <- ces2019_phone %>% 
  select(7, 8, 9, 20, 25:30)

@@

We can see that some of the column names of the subset are not clearly named. The remedy this we can use the `rename` function from the `dplyr` package (Wickham et al., 2020). When using `rename()`, the order goes from new name to old name (see [*3.1.6: Rename columns*](#316-rename-columns)). To gain an idea of what to rename a column we can use the `get_question()` function to find the associated survey question (see [*3.1.5: `get_question()`*](#315-get_question)).

#### `get_question()`
```
get_question("ces2019_phone_subset", "q1")
get_question("ces2019_phone_subset", "q2")
get_question("ces2019_phone_subset", "q3")
get_question("ces2019_phone_subset", "q4")
get_question("ces2019_phone_subset", "q6")
get_question("ces2019_phone_subset", "q7")
```
```
Are you a Canadian Citizen?
In what year were you born?
Gender
In which province or territory are you currently living?
On the whole, are you very satisfied, fairly satisfied, not very satisfied
Most important issue in this FEDERAL election
```

Using the labels returned by `get_question()` we can better name the subset columns (see [*3.1.6: Rename colums*](#316-rename-columns)).

#### Rename columns
```
ces2019_phone_subset <- ces_phone_subset %>% 
  rename(interviewer_gender = interviewer_gender_CES,
         language = language_CES,
         phonetype = phonetype_CES,
         weight = weight_CES,
         citizenship = q1,
         yob = q2,
         respondent_gender = q3,
         province_territory = q4,
         dmcrcy_satisfaction = q6,
         imprtnt_issue = q7)
```
This will have now created a subset of the CES 2019 phone survey with renamed variables and factor values.

>>>>

When called, the `get_ces()` function returns a requested CES survey as a data object and prints to the console the associated citation and URL for the survey dataset repository. The function takes one argument in the form of a character string. This argument is a vector member that has been associated with a CES survey through the body of code in the `get_ces()` function that when used calls the download URL for that survey on an associated GitHub repository named `ces_data`. If the provided character string argument matches a member of the built-in vector `ces_codes`, the associated file is downloaded using the `download.file()` function from the `utils` R package (R Core Team, 2020) as a compressed .zip folder and is stored temporarily in `inst/extdata` directory in the greater package directory. Upon downloading the file, the compressed folder is unzipped using the `unzip()` function from the `utils` R package (R Core Team, 2020) and read into R using either the `read_dta()` or `read_sav()` functions from the `haven` R package [@citehaven] depending on the file extension of the downloaded file. A data frame is then assigned using the `assign()` function  from the `base` R package (R Core Team, 2020) as a data object in the global environment. The downloaded file and file directory are then removed from the computer using the `unlink()` function from the `base` R package (R Core Team, 2020). Finally, the recommended citation for the requested survey dataset and URL of the survey data storage location are printed in the console (see [*2.1.1 `get_ces()` basics*](#211-get_ces-basics) for an example of a basic `get_ces()` function call).

If the provided character string argument does not have a match in the built-in vector, then the function process is stopped and a warning message stating `Error in get_ces(): Warning: Code not in table` is printed in the RStudio console (see [*2.1.2 `get_ces()` error*](#212-get_ces-error)).

### `get_ces()` basics
```
# install the cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

### `get_ces()` error
```
# install the cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

CES survey dataset files uses either a `.dta()` or `.sav()` file extension, meaning the datasets are loaded into R and RStudio as the type labelled. The `get_ces()` function does not convert the values of the loaded tables to a factor type so that personal workflow practices are not interfered with. It is suggested that to convert the dataset values to a factor type that the `to_factor()` function from the `labelled` package (Lamarange, 2020) be used (see [*2.1.3 `to_factor()` function*](#213-to_factor-function) for an example as to using the `to_factor()` function and `labelled` package).

### `to_factor()` function
```
# install cesR package from GitHub
devtools::intall_github("hodgettsp/cesR")

@@

The `get_cescodes()` function provides a user a means of looking up the argument calls that are used to access each CES survey dataset. The `get_cescodes()` function does not take any arguments. Instead when the function is called it prints to the console a dataframe that contains the survey codes and their associated argument calls. See [*2.2.1 `get_cescodes()` function*](#221-get_cescodes-function) for an example of the print output and the [*3. Vignette*](#3-vignette) section for an example of the `get_cescodes()` used as part of the full `cesR` package.

### `get_cescodes()` function
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

## `get_question()`
The `get_question()` provides users with the ability to look up a survey question associated with a given column name. The function takes two arguments in the form of character strings, those being the name of a data object and the name of a column in the given data object. The function works such that it checks whether the given data object exists using the `exists()` function from the `base` package (R Core Team, 2020). If the object does not exist, the function will print out a warning in the console stating `Warning: Data object does not exist` (see [*2.3.2 `get_question()` data object error*](#232-get_question-data-object-error) for an example of this error). If the object does exist, `get_question()` will check if the given column name exists in the given data object. This is done using a combination of the `hasName()` function from the `utils` package and the `get()` function from the `base` package (R Core Team, 2020). The `hasName()` function checks if the given column name is in the given data object. Because the arguments are given as character strings the `get()` function is used to return the actual data object instead of the provided character string. Otherwise, the `hasName()` function would only check if the given column name argument occurred in the given character string argument and not the actual data object. If the column does not exist in the data object a warning is printed in the console stating `Warning: Variable is not in dataset`. See [*2.3.3 `get_question()` variable error*](#233-get_question-variable-error) for an example of this error message. If the given column exists in the given data object, `get_question()` will print the variable label of the given column to the console using a combination of the `var_label()` function from the `labelled` package (Larmarange, 2020) and the `get()` function from the `base` package (R Core Team, 2020). See [*2.3.1 `get_question()` basics*](#231-get_question-basics) for an example of the general use of the `get_question()` function and the [*3. Vignette*](#3-vignette) section for its use as part of the whole `cesR` package.

### `get_question()` basics
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

### `get_question()` data object error
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

### `get_question()` variable error
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

In addition to being usable with the data objects created from the [`get_ces()`](#21-get_ces) function, the `get_question()` function is structured in such a way that it can also be used to return the column label for the `decon` dataset or any dataset of the labelled type. [*2.3.4 `get_question()` `decon`*](#234-get_question-decon) presents an example of the `get_question()` function being used with the `decon` dataset.

### `get_question()` `decon`
```
# install cesR package from GitHub
devtools::install_github("hodgettsp/cesR")

@@

## `get_preview()`

To make the data exploratory phase easier it may be prudent to have access to a truncated version of a dataset. Additionally, a truncated dataset provides a resource that can be used by educators in the teaching of exploratory data analysis. The `get_preview()` function provides such truncated versions of the CES datasets. The function takes two arguments, a character string to call a survey of the same style as used for the [`get_ces()`](#21-get_ces) function and a numerical value that sets the number of rows returned (see [2.4.1 `get_preview()` basics](#241-get_preview-basics)). If no value is provided for the number of rows, a default of six is returned (see [2.4.2 `get_preview()` no value provided](#242-get_preview-no-value-provided)).

### `get_preview()` basics
```
#install cesR package
devtools::install_github("hodgettsp/cesR")

@@

### `get_preview()` no value provided
```
# install cesR package
devtools::install_github("hodgettsp/cesR")

@@

## get_decon()
When called, creates a subset of the 2019 CES online survey under the name `decon` (demographics and economics) that provides a tool for educators in the teaching of the analysis of large survey datasets. The `get_decon()` function takes no arguments. The function first checks the global environment if an object named `decon` exists using the `exists()` function from the `base` package (R Core Team, 2020). This prevents the `decon` dataset from being recreated if the object exists. If the `get_decon()` function is run when an object with the name `decon` already exists a warning will print in the console stating `Error in get_decon() : Warning: File already exists.` See [*2.5.1 `get_decon()` error*](#251-get_decon-error) for an example of the error message. If a situation arises in which the `decon` dataset needs to be recreated, then the best course of action is to use the `rm()` function from the `base` package (R Core Team, 2020) to remove the `decon` object and then run the `get_decon()` function again (see [*2.5.2 `get_decon()` reload*](#252-get_decon-reload)).

### `get_decon()` error
```
# install cesR package
devtools::install_github("hodgettsp/cesR")

@@

### `get_decon()` reload
```
# install cesR package
devtools::install_github("hodgettsp/cesR")

@@

After checking if an object with the name `decon` exists, the `get_decon()` function performs similarly to the `get_ces()` function. It assigns a temporary file extension with a .zip type using the `tempfile()` function from the `base` package, and calls on the URL for the 2019 CES online survey and temporarily downloads the associated file using the `download.file()` function from the `base` package (R Core Team, 2020). After unpacking the compressed folder with the `unzip()` function from the `utils` package (R Core Team, 2020), the file is read into R using the `read_dta()` function from the `haven` package (Wickham & Miller, 2020). Using the `select()` function from the `dplyr` package, 20 variables are selected from the main CES 2019 online survey and renamed using the `rename` function from the `dplyr` package (Wickham et al., 2020). The dataframe values are then converted to a factor type using the `to_factor()` function from the `labelled` package (Lamarange, 2020). A new variable, consisting of participants' responses to their self-perceived position on the political spectrum, is then created from two variables using the `unite()` function from the `tidyr` package (Wickham & Henry, 2020) and all empty cells are replaced with `NA` values using the `mutate()` function from the `dplyr` package (Wickham et al., 2020). The dataset is made available in the global environment by using the `assign()` function from the `base` package to create a new data object under the name `decon`. Lastly, the downloaded files are removed from the computer using the `unlink()` function from the `base` package (R Core Team, 2020) and a citation for the 2019 CES online survey and a URL to the storage location is printed to the console. See [*2.5.3 `get_decon()` basics*](#253-get_decon-basics) for an example of a general use of the `get_decon()` function.

### `get_decon()` basics

@@

To begin, install and load the `cesR` package (and all other necessary packages) into RStudio. Currently, this is currently only available through the use of the `install_github` function from the `devtools` package (Wickham et al., 2020). During installation, RStudio may request to update other packages. It is best to press enter with an empty line (see [*3.1.1 Install `cesR`*](#311-install-cesR)).

### Install `cesR`
```
# uncomment any package that needs to be installed
# install.packages("devtools")
# install.packages("labelled")
# install.packages("dplyr")
# install.packages("tidyr")

@@

Upon installation and loading of the `cesR` package, we can use the `get_cescodes()` function to look up the code for the desired CES survey. In this case that is the 2019 CES phone survey (see [*3.1.2 Lookup CES codes*](#312-lookup-ces-codes)).

### Lookup CES codes
```
# look up survey codes
get_cescodes()
```
```
> get_cescodes()
   index ces_survey_code get_ces_call_char
1      1     ces2019_web     "ces2019_web"
2      2   ces2019_phone   "ces2019_phone"
3      3     ces2015_web     "ces2015_web"
4      4   ces2015_phone   "ces2015_phone"
5      5   ces2015_combo   "ces2015_combo"
6      6         ces2011         "ces2011"
7      7         ces2008         "ces2008"
8      8         ces2004         "ces2004"
9      9         ces0411         "ces0411"
10    10         ces0406         "ces0406"
11    11         ces2000         "ces2000"
12    12         ces1997         "ces1997"
13    13         ces1993         "ces1993"
14    14         ces1988         "ces1988"
15    15         ces1984         "ces1984"
16    16         ces1974         "ces1974"
17    17         ces7480         "ces7480"
18    18      ces72_jnjl      "ces72_jnjl"
19    19       ces72_sep       "ces72_sep"
20    20       ces72_nov       "ces72_nov"
21    21         ces1968         "ces1968"
22    22         ces1965         "ces1965"
```
In the printed data frame we can see that the argument code for the 2019 CES phone survey is "ces2019_phone" (see [*3.1.2 Lookup CES codes*](#312-lookup-ces-codes)). This code can be used with the `get_preview()` function to look up a truncated preview of the 2019 CES phone survey dataset (see [*3.1.3 Preview dataset*](#313-preview-dataset)).

### Preview dataset
```
# get preview of first 8 rows of 2019 CES phone survey dataset
get_preview("ces2019_phone", 8)
```
 sample_id | survey_end_CES | survey_end_mont~ | survey_end_day_~ | num_attempts_CES | interviewer_id_~ | interviewer_gen~ | language_CES | phonetype_CES |
-----------|----------------|------------------|------------------|------------------|------------------|------------------|--------------|---------------|
<dbl>      | <chr>          | <dbl>            | <dbl>            |  <dbl>           |  <dbl>           |     <chr>        |    <fct>     | <fct>         |
18         | 2019-09-23 15~ |               9  |              23  |              5   |         161182   | Female           | (2) French   |  (2) Wireless |
32         | 2019-09-12 18~ |               9  |              12  |              1   |         151152   | Male             | (1) English  |  (2) Wireless |
39         | 2019-09-10 18~ |               9  |              10  |              1   |         161182   | Female           | (2) French   |  (2) Wireless |
59         | 2019-10-10 15~ |              10  |              10  |              6   |         147601   | Female           | (2) French   |  (2) Wireless |
61         | 2019-09-12 16~ |               9  |              12  |              1   |         151152   | Male             | (2) French   |  (2) Wireless |
69         | 2019-09-17 17~ |               9  |              17  |              1   |           2503   | Female           | (2) French   |  (2) Wireless |       
157        | 2019-09-12 16~ |               9  |              12  |              1   |           2529   | Male             |  (2) French  | (2) Wireless   | 
158        | 2019-09-14 10~ |                9 |              14  |               4  |         142515   | Male             |  (2) French  | (2) Wireless   |

Using the same survey code argument with the `get_ces()` function we can retrieve the complete 2019 CES phone survey (see [*3.1.4 `get_ces()`*](#314-get_ces)). Additionally, we can see when `get_ces()` is called, the survey citation and link to repository prints to the console.

### `get_ces()`
```
# call 2019 CES phone survey
get_ces("ces2019_phone")
```
```
TO CITE THIS SURVEY FILE: Stephenson, Laura B; Harell, Allison; Rubenson, Daniel; Loewen, Peter John, 2020, '2019 Canadian Election Study - Phone Survey', https://doi.org/10.7910/DVN/8RHLG1, Harvard Dataverse, V1, UNF:6:eyR28qaoYlHj9qwPWZmmVQ== [fileUNF]

@@

To check that the dataset has loaded into RStudio correctly, it is best to check the dataset with the `head()` function from the `utils` package (R Core Team, 2020). Unfortunately, the `head()` function does not work with labelled data, so it is best to convert the values to factors using the `to_factor()` function from the `labelled` package (Larmarange, 2020) (see [*3.1.5 Check 2019 CES phone data*](#315-check-ces-phone-data)).

### Check 2019 CES phone data
```
# convert values to factor type
ces2019_phone <- to_factor(ces2019_phone)
head(ces2019_phone)

@@

With the dataset loaded, the `select()` function from the `dplyr` package (Wickham et al., 2020) is an efficient way of selecting the required variable columns. In this case, the language, phone type, weight, citizenship, year of birth, interviewer gender, identified gender, province, satisfaction with Canadian democracy, and most important election issue are selected using the column index (see [*3.1.4 Select columns*](#314-select-columns)). Additionally, the columns are not all named in an understandable way and so will be renamed using the `rename()` function from the `dplyr` package (Wickham et al., 2020).

### Select columns
```
ces2019_phone_subset <- ces2019_phone %>% 
  select(7, 8, 9, 20, 25:30)

@@

We can see that some of the column names of the subset are not clearly named. The remedy this we can use the `rename` function from the `dplyr` package (Wickham et al., 2020). When using `rename()`, the order goes from new name to old name (see [*3.1.6 Rename columns*](#316-rename-columns)). To gain an idea of what to rename a column we can use the `get_question()` function to find the associated survey question (see [*3.1.5 `get_question()`*](#315-get_question)).

### `get_question()`
```
get_question("ces2019_phone_subset", "q1")
get_question("ces2019_phone_subset", "q2")
get_question("ces2019_phone_subset", "q3")
get_question("ces2019_phone_subset", "q4")
get_question("ces2019_phone_subset", "q6")
get_question("ces2019_phone_subset", "q7")
```
```
Are you a Canadian Citizen?
In what year were you born?
Gender
In which province or territory are you currently living?
On the whole, are you very satisfied, fairly satisfied, not very satisfied
Most important issue in this FEDERAL election
```

Using the labels returned by `get_question()` we can better name the subset columns (see [*3.1.6 Rename colums*](#316-rename-columns)).

### Rename columns
```
ces2019_phone_subset <- ces_phone_subset %>% 
  rename(interviewer_gender = interviewer_gender_CES,
         language = language_CES,
         phonetype = phonetype_CES,
         weight = weight_CES,
         citizenship = q1,
         yob = q2,
         respondent_gender = q3,
         province_territory = q4,
         dmcrcy_satisfaction = q6,
         imprtnt_issue = q7)
```
This will have now created a subset of the CES 2019 phone survey with renamed variables and factor values.
